---
title: "Reproducible Research: Peer Assessment 1"
author: "Thomas Kanold"
date: "15. November 2014"
output: html_document
---


## Loading and preprocessing the data
```{r libraries,message=FALSE, cache=FALSE}
library(dplyr)
```

```{r loading data}


fileName <- "activity.csv"
surv <- read.csv(fileName)
surv  <- tbl_df(surv)
```


## What is mean total number of steps taken per day?
```{r}
steps_taken <- surv %>% 
                select(steps, date) %>% 
                filter(!is.na(steps)) %>% 
                group_by(date) %>%  
                summarize(mean(steps))
st <- steps_taken[,2]
```

```{r histogram}
hist(st, breaks=20, freq=FALSE, xlab="steps taken", main="Daily Steps", col="lightgreen")
#curve(dnorm(x, mean=mean(st), sd=sd(st)), add=TRUE, col="darkblue", lwd=2)


st_mean <- mean(st)
st_median <- median(st)
```

The mean total number of steps taken per day is `r round(st_mean,3)`, the median is `r round(st_median,3)`.



## What is the average daily activity pattern?

```{r plotting average steps a day}

steps_taken <- surv %>% 
  select(steps, interval) %>% 
  filter(!is.na(steps)) %>% 
  group_by(interval) %>%  
  summarize(mm = mean(steps))
st5 <- steps_taken[,1:2]

plot(st5, type='l', main="Steps a Day")
```

```{r max_average}
max_steps <- surv %>% 
  select(steps, interval) %>% 
  filter(!is.na(steps)) %>% 
  group_by(interval) %>%  
  summarize(mm = mean(steps))%>% 
  arrange(desc(mm) )

```


The time interval `r max_steps[1,1]` has the highest average (over all days)  step count with `r round(max_steps[1,2],3)`.


## Imputing missing values


```{r}
na_steps <- surv %>% filter(is.na(steps)) %>% 
  select(steps, interval) 
```

The number of rows with missing values is `r nrow(na_steps)`.

The strategy to fill in the missing NA values is choosing the average value of that time interval.
```{r NA to something}
# A function that takes the input and gives it back 
# NA will be the average of this time interval
filler <- function (steps, mm) { 
  res <- ifelse(!is.na(steps),steps,mm)
}

survm <- merge(surv, st5)
survm <- mutate(survm, steps = filler(steps, mm))

survm <- survm %>% select(interval, date, steps) %>% arrange(desc(date))

```

```{r}
steps_taken <- survm %>% select(steps, date) %>% filter(!is.na(steps)) %>% group_by(date) %>%  summarize(mean(steps))
st <- steps_taken[,2]
hist(st, breaks=20, freq=FALSE, xlab="steps taken", main="Daily Steps", col="lightgreen")
#curve(dnorm(x, mean=mean(st), sd=sd(st)), add=TRUE, col="darkblue", lwd=2)


survm_mean <- mean(st)
survm_median <- round(median(st))
```

The mean total number of steps taken per day is `r round(survm_mean,3)`, 
the median is `r round(survm_median,3)`.
The difference between the mean values is `r survm_mean - st_mean`, median difference is
`r survm_median - st_median`.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
####


itsAWeekend <- function(date) {
    date2 <- as.POSIXlt(surv$date)$wday
    #weekdays(as.Date(surv$date))
    res <- ifelse(date2==0 | date2==6, "weekend", "weekday")  
}

survm_weekends <- mutate(survm, dayType = itsAWeekend(date))



max_steps <- survm_weekends %>% 
  select(steps, interval, dayType) %>% 
  group_by(dayType, interval) %>%  
  summarize(mm = mean(steps))


par(mfrow=c(2,1)) 

plot(max_steps[max_steps$dayType == "weekday",2:3], type='l', main="weekday")
plot(max_steps[max_steps$dayType != "weekday",2:3], type='l', main="weekend")

```









